---
title: "CompareMaps documentation"
output: 
  github_document:
    pandoc_args: [
      "--output=README.md"
    ]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# NAME


CompareMaps - It's an R script to compare two phylostratigraphic maps and visualize this comparison using heatmap.


# SYNOPSIS

```{r, echo=TRUE, eval=FALSE}

CompareMaps.R --map1 file --map2 file --out outdir

CompareMaps.R --map1 ./t/data/bsfull3.phmap_names --map2 ./t/data/bscdhit3.phmap_names --out ./t/data/
```

# DESCRIPTION
This R script takes as input two maps generated by PhyloStrat software. That means they have headers with phylogeny information + stratified genes in the body of document.
It uses data.table to calculate joins between all phylostrata of one map and all phylostrata of other map.
After that it plots this grid onto a heatmap and it prints this plot into pdf and png files.
It also writes map comparison into .tsv, .xlsx and .Rdata files.

```{r, echo=FALSE, eval=TRUE}
my_heatmap <- function (x, Rowv = NULL, Colv = if (symm) "Rowv" else NULL, 
    distfun = dist, hclustfun = hclust, reorderfun = function(d, 
        w) reorder(d, w), add.expr, symm = FALSE, revC = TRUE, scale = c("row", "column", "none"), na.rm = TRUE, 
    margins = c(5, 5), ColSideColors, RowSideColors, cexRow = 0.2 + 
        1/log10(nr), cexCol = 0.2 + 1/log10(nc), labRow = NULL, 
    labCol = NULL, main = NULL, xlab = NULL, ylab = NULL, keep.dendro = FALSE, 
    verbose = getOption("verbose"), ...) 
{
    # create copy for printing values
    my_org <- x
    
    #normal
    scale <- if (symm && missing(scale)) 
        "none"
    else match.arg(scale)
    if (length(di <- dim(x)) != 2 || !is.numeric(x)) 
        stop("'x' must be a numeric matrix")
    nr <- di[1L]
    nc <- di[2L]
    if (nr <= 1 || nc <= 1) 
        stop("'x' must have at least 2 rows and 2 columns")
    if (!is.numeric(margins) || length(margins) != 2L) 
        stop("'margins' must be a numeric vector of length 2")
    doRdend <- !identical(Rowv, NA)
    doCdend <- !identical(Colv, NA)
    if (!doRdend && identical(Colv, "Rowv")) 
        doCdend <- FALSE
    if (is.null(Rowv)) 
        Rowv <- rowMeans(x, na.rm = na.rm)
    if (is.null(Colv)) 
        Colv <- colMeans(x, na.rm = na.rm)
    if (doRdend) {
        if (inherits(Rowv, "dendrogram")) 
            ddr <- Rowv
        else {
            hcr <- hclustfun(distfun(x))
            ddr <- as.dendrogram(hcr)
            if (!is.logical(Rowv) || Rowv) 
                ddr <- reorderfun(ddr, Rowv)
        }
        if (nr != length(rowInd <- order.dendrogram(ddr))) 
            stop("row dendrogram ordering gave index of wrong length")
    }
    else rowInd <- 1L:nr
    if (doCdend) {
        if (inherits(Colv, "dendrogram")) 
            ddc <- Colv
        else if (identical(Colv, "Rowv")) {
            if (nr != nc) 
                stop("Colv = \"Rowv\" but nrow(x) != ncol(x)")
            ddc <- ddr
        }
        else {
            hcc <- hclustfun(distfun(if (symm) 
                x
            else t(x)))
            ddc <- as.dendrogram(hcc)
            if (!is.logical(Colv) || Colv) 
                ddc <- reorderfun(ddc, Colv)
        }
        if (nc != length(colInd <- order.dendrogram(ddc))) 
            stop("column dendrogram ordering gave index of wrong length")
    }
    else colInd <- 1L:nc
    x <- x[rowInd, colInd]
    labRow <- if (is.null(labRow)) 
        if (is.null(rownames(x))) 
            (1L:nr)[rowInd]
        else rownames(x)
    else labRow[rowInd]
    labCol <- if (is.null(labCol)) 
        if (is.null(colnames(x))) 
            (1L:nc)[colInd]
        else colnames(x)
    else labCol[colInd]
    
    # changed to color relative values
    if (scale == "row") {
        maxx <- suppressWarnings(apply(x, 1L, max, na.rm = T))
        x <- sweep(x, 1L, maxx, "/", check.margin = FALSE)
    }
    # changed to color relative values
    else if (scale == "column") {
        maxx <- suppressWarnings(apply(x, 2L, max, na.rm = T))
        x <- sweep(x, 2L, maxx, "/", check.margin = FALSE)
    }
    lmat <- rbind(c(NA, 3), 2:1)
    lwid <- c(if (doRdend) 1 else 0.05, 4)
    lhei <- c((if (doCdend) 1 else 1) + if (!is.null(main)) 0.2 else 0, 
        4)
    if (!missing(ColSideColors)) {
        if (!is.character(ColSideColors) || length(ColSideColors) != 
            nc) 
            stop("'ColSideColors' must be a character vector of length ncol(x)")
        lmat <- rbind(lmat[1, ] + 1, c(NA, 1), lmat[2, ] + 1)
        lhei <- c(lhei[1L], 0.2, lhei[2L])
    }
    if (!missing(RowSideColors)) {
        if (!is.character(RowSideColors) || length(RowSideColors) != 
            nr) 
            stop("'RowSideColors' must be a character vector of length nrow(x)")
        lmat <- cbind(lmat[, 1] + 1, c(rep(NA, nrow(lmat) - 1), 
            1), lmat[, 2] + 1)
        lwid <- c(lwid[1L], 0.2, lwid[2L])
    }
    lmat[is.na(lmat)] <- 0
    if (verbose) {
        cat("layout: widths = ", lwid, ", heights = ", lhei, 
            "; lmat=\n")
        print(lmat)
    }
    dev.hold()
    on.exit(dev.flush())
    op <- par(no.readonly = TRUE)
    on.exit(par(op), add = TRUE)
    layout(lmat, widths = lwid, heights = lhei, respect = TRUE)
    if (!missing(RowSideColors)) {
        par(mar = c(margins[1L], 0, 0, 0.5))
        image(rbind(if (revC) 
            nr:1L
        else 1L:nr), col = RowSideColors[rowInd], axes = FALSE)
    }
    if (!missing(ColSideColors)) {
        par(mar = c(0.5, 0, 0, margins[2L]))
        image(cbind(1L:nc), col = ColSideColors[colInd], axes = FALSE)
    }
    par(mar = c(margins[1L], 0, 0, margins[2L]))
    if (!symm || scale != "none") 
        x <- t(x)
    if (revC) {
        iy <- nr:1
        if (doRdend) 
            ddr <- rev(ddr)
        x <- x[, iy]
        # switch rows
        my_org <- my_org[iy,]
    }
    else iy <- 1L:nr
    image(1L:nc, 1L:nr, x, xlim = 0.5 + c(0, nc), ylim = 0.5 + 
        c(0, nr), axes = FALSE, xlab = "", ylab = "", ...)
    #changed 1 to 3
    axis(3, 1L:nc, labels = labCol, las = 2, line = -0.5, tick = 0, 
        cex.axis = cexCol)
    if (!is.null(xlab)) 
        mtext(xlab, side = 3, line = margins[1L] )
    axis(2, iy, labels = labRow, las = 2, line = -0.5, tick = 0, 
        cex.axis = cexRow)
    
    # printing of values
    for (xcord in 1:ncol(my_org)) {
        for (ycord in 1:nrow(my_org)) {
            if (is.na(my_org[ycord,xcord])) {
                text(xcord, ycord, '-')
            }
            else {
                text(xcord, ycord, my_org[ycord,xcord])
            }
        }
    }
            
    # try printing box around heatmap
    box(lty="solid", col="black")
    
    # try printing grid around values
    grid(nrow(my_grid), ncol(my_grid))
    
    #changed 4 to 2
    if (!is.null(ylab)) 
        mtext(ylab, side = 2, line = margins[2L] )
    if (!missing(add.expr)) 
        eval.parent(substitute(add.expr))
    par(mar = c(margins[1L], 0, 0, 0))
    if (doRdend) 
        plot(ddr, horiz = TRUE, axes = FALSE, yaxs = "i", leaflab = "none")
    else frame()
    par(mar = c(0, 0, if (!is.null(main)) 1 else 0, margins[2L]))
    if (doCdend) 
        plot(ddc, axes = FALSE, xaxs = "i", leaflab = "none")
    else if (!is.null(main)) 
        frame()
    if (!is.null(main)) {
        par(xpd = NA)
        title(main, cex.main = 1.5 * op[["cex.main"]])
    }
    invisible(list(rowInd = rowInd, colInd = colInd, Rowv = if (keep.dendro && 
        doRdend) ddr, Colv = if (keep.dendro && doCdend) ddc))
}
```


```{r, echo=TRUE, fig.width=10, fig.height=10}
load(file="./t/data/my_grid.Rdata")
my_heatmap(my_grid, Colv = NA, Rowv = NA, scale="row", col=rev(heat.colors(20)),
    main=paste("bsfull3", "(1)_vs_", "bscdhit3", "(2)_comparison", sep=""), xlab="bscdhit3", ylab="bsfull3")
```

# AUTHOR

Martin Sebastijan Šestak (<martin.s.sestak@gmail.com>)

# LICENCE AND COPYRIGHT

Copyright (c) 2016-2017 Martin Sebastijan Šestak. All rights reserved.

This module is free software; you can redistribute it and/or
modify it under the GPL version 3.

This program is distributed in the hope that it will be useful,
but WITHOUT ANY WARRANTY; without even the implied warranty of
MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. 
