---
title: "CompareMaps_work_in_progress"
author: "Martin Sebastijan Šestak"
date: "12/9/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## CompareMaps workflow from RStudio

This is created so I don't need to rerun code every time from scratch.
```{r, echo=TRUE}
library("data.table")
library("sqldf")

```


```{r, echo=TRUE}
getwd()
setwd("/msestak/gitdir/CompareMaps/t/data")
getwd()
list.files()
```

get phylogeny from one of maps

```{r, echo=TRUE}
phylogeny <- read.table(text=readLines("bsfull3.phmap_names")[count.fields("bsfull3.phmap_names", sep="\t")==3], sep="\t", stringsAsFactors=F)
phylogeny <- phylogeny[-1,]
colnames(phylogeny) = c("ps", "psti", "psname")
phylogeny$psname <- sub("[^:]+:\\s+", "", phylogeny$psname, perl=T)
phylogeny
```

get map content from files
```{r, echo=TRUE}
map1 <- read.table(text=readLines("bsfull3.phmap_names")[count.fields("bsfull3.phmap_names", sep="\t")==4], sep="\t", stringsAsFactors=F)
map2 <- read.table(text=readLines("bscdhit3.phmap_names")[count.fields("bscdhit3.phmap_names", sep="\t")==4], sep="\t", stringsAsFactors=F)
colnames(map1) = c("prot_id", "ps", "psti", "psname")
map1$psname <- sub("[^:]+:\\s+", "", map1$psname, perl=T)
colnames(map2) = c("prot_id", "ps", "psti", "psname")
map2$psname <- sub("[^:]+:\\s+", "", map2$psname, perl=T)
head(map1)
```
create a grid to be used later to store results of joins
```{r, echo=TRUE}
my_grid <- matrix(data=NA, nrow=nrow(phylogeny), ncol=nrow(phylogeny))
rownames(my_grid) <- paste("map1_ps", 1:nrow(my_grid), sep="")
colnames(my_grid) <- paste("map2_ps", 1:ncol(my_grid), sep="")
my_grid
```
# convert data.frames to data.table to use data.table joins on them
```{r, ECHO=TRUE}
mapa1 = as.data.table(map1)
mapa2 = as.data.table(map2)
setkey(mapa1,prot_id)
setkey(mapa2,prot_id)
system.time(ps1 <- mapa1[mapa2, .N, nomatch=0, on = "prot_id"])
ps1 <- mapa1[mapa2, NROW(prot_id), nomatch=0, on = "prot_id"]
system.time(ps1 <- mapa1[mapa1$ps==1][mapa2[mapa2$ps==1], .N, nomatch=0, on = "prot_id"])
ps1

```
do the joins in a double loop to iterate over all positions in a matrix
```{r, echo=TRUE}
for (map1 in phylogeny$ps) {
    for (map2 in phylogeny$ps) {
        #print(my_grid[map1,map2])
        my_grid[map1,map2] <- mapa1[mapa1$ps==map1][mapa2[mapa2$ps==map2], NROW(prot_id), nomatch=0, on = "prot_id"]
        print(my_grid[map1,map2])
    }
}

```

create a levelplot
```{r, eval=F, echo=T, warning=FALSE}
M <- cor(my_grid, method = "pearson")
levelplot(M, col.regions=rev(heat.colors(16)))

```


```{r, eval=F, echo=TRUE}
my_colors <-colorRampPalette(brewer.pal(9,"Reds"))(16)
levelplot(M, col.regions=my_colors)

z <- as.matrix(dist(my_grid, method="canberra"))
levelplot(z, col.regions=rev(my_colors))

dd <- as.matrix(as.dist((1-cor(my_grid))/2))
levelplot(dd, col.regions=rev(my_colors))

levelplot(my_grid, col.regions=rev(gray(0:100/100)), xlab = "X koordinata", ylab = "Y koordinata", main="Naslov", scale=list(x=list(rot=45)))

heatmap(my_grid, Colv = NA, Rowv = NA, scale="row", col=rev(gray(0:100/100)))

my_grid[my_grid < 1] <- NA
my_heatmap3(my_grid, Colv = NA, Rowv = NA, scale="row", col=my_colors, main="ma1_vs_map2_comparison")
my_heatmap16(my_grid, Colv = NA, Rowv = NA, scale="row", col=rev(heat.colors(100)), main="ma1_vs_map2_comparison")
my_heatmap19(my_grid, Colv = NA, Rowv = NA, scale="row", col=rev(heat.colors(20)), main="map1_vs_map2_comparison", xlab="map2", ylab="map1")
```

my_heatmap(my_grid, Colv = NA, Rowv = NA, scale="row", col=rev(heat.colors(20)),
    main=paste(map1_name, "_vs_", map2_name, "_comparison", sep=""), xlab=map2_name, ylab=paste(map1_name, sep=""))

write to table
```{r, echo=TRUE}
write.table(my_grid, file="my_grid", sep="\t")
save(my_grid, file="my_grid.Rdata")

```

show heatmap code
```{r, echo=T}
body(heatmap)

```

show my_heatmap19 code
```{r, echo=F}
function (x, Rowv = NULL, Colv = if (symm) "Rowv" else NULL, 
    distfun = dist, hclustfun = hclust, reorderfun = function(d, 
        w) reorder(d, w), add.expr, symm = FALSE, revC = TRUE, scale = c("row", "column", "none"), na.rm = TRUE, 
    margins = c(5, 5), ColSideColors, RowSideColors, cexRow = 0.2 + 
        1/log10(nr), cexCol = 0.2 + 1/log10(nc), labRow = NULL, 
    labCol = NULL, main = NULL, xlab = NULL, ylab = NULL, keep.dendro = FALSE, 
    verbose = getOption("verbose"), ...) 
{
    # create copy for printing values
    my_org <- x
    
    #normal
    scale <- if (symm && missing(scale)) 
        "none"
    else match.arg(scale)
    if (length(di <- dim(x)) != 2 || !is.numeric(x)) 
        stop("'x' must be a numeric matrix")
    nr <- di[1L]
    nc <- di[2L]
    if (nr <= 1 || nc <= 1) 
        stop("'x' must have at least 2 rows and 2 columns")
    if (!is.numeric(margins) || length(margins) != 2L) 
        stop("'margins' must be a numeric vector of length 2")
    doRdend <- !identical(Rowv, NA)
    doCdend <- !identical(Colv, NA)
    if (!doRdend && identical(Colv, "Rowv")) 
        doCdend <- FALSE
    if (is.null(Rowv)) 
        Rowv <- rowMeans(x, na.rm = na.rm)
    if (is.null(Colv)) 
        Colv <- colMeans(x, na.rm = na.rm)
    if (doRdend) {
        if (inherits(Rowv, "dendrogram")) 
            ddr <- Rowv
        else {
            hcr <- hclustfun(distfun(x))
            ddr <- as.dendrogram(hcr)
            if (!is.logical(Rowv) || Rowv) 
                ddr <- reorderfun(ddr, Rowv)
        }
        if (nr != length(rowInd <- order.dendrogram(ddr))) 
            stop("row dendrogram ordering gave index of wrong length")
    }
    else rowInd <- 1L:nr
    if (doCdend) {
        if (inherits(Colv, "dendrogram")) 
            ddc <- Colv
        else if (identical(Colv, "Rowv")) {
            if (nr != nc) 
                stop("Colv = \"Rowv\" but nrow(x) != ncol(x)")
            ddc <- ddr
        }
        else {
            hcc <- hclustfun(distfun(if (symm) 
                x
            else t(x)))
            ddc <- as.dendrogram(hcc)
            if (!is.logical(Colv) || Colv) 
                ddc <- reorderfun(ddc, Colv)
        }
        if (nc != length(colInd <- order.dendrogram(ddc))) 
            stop("column dendrogram ordering gave index of wrong length")
    }
    else colInd <- 1L:nc
    x <- x[rowInd, colInd]
    labRow <- if (is.null(labRow)) 
        if (is.null(rownames(x))) 
            (1L:nr)[rowInd]
        else rownames(x)
    else labRow[rowInd]
    labCol <- if (is.null(labCol)) 
        if (is.null(colnames(x))) 
            (1L:nc)[colInd]
        else colnames(x)
    else labCol[colInd]
    
    # changed to color relative values
    if (scale == "row") {
        maxx <- suppressWarnings(apply(x, 1L, max, na.rm = T))
        x <- sweep(x, 1L, maxx, "/", check.margin = FALSE)
    }
    # changed to color relative values
    else if (scale == "column") {
        maxx <- suppressWarnings(apply(x, 2L, max, na.rm = T))
        x <- sweep(x, 2L, maxx, "/", check.margin = FALSE)
    }
    lmat <- rbind(c(NA, 3), 2:1)
    lwid <- c(if (doRdend) 1 else 0.05, 4)
    lhei <- c((if (doCdend) 1 else 1) + if (!is.null(main)) 0.2 else 0, 
        4)
    if (!missing(ColSideColors)) {
        if (!is.character(ColSideColors) || length(ColSideColors) != 
            nc) 
            stop("'ColSideColors' must be a character vector of length ncol(x)")
        lmat <- rbind(lmat[1, ] + 1, c(NA, 1), lmat[2, ] + 1)
        lhei <- c(lhei[1L], 0.2, lhei[2L])
    }
    if (!missing(RowSideColors)) {
        if (!is.character(RowSideColors) || length(RowSideColors) != 
            nr) 
            stop("'RowSideColors' must be a character vector of length nrow(x)")
        lmat <- cbind(lmat[, 1] + 1, c(rep(NA, nrow(lmat) - 1), 
            1), lmat[, 2] + 1)
        lwid <- c(lwid[1L], 0.2, lwid[2L])
    }
    lmat[is.na(lmat)] <- 0
    if (verbose) {
        cat("layout: widths = ", lwid, ", heights = ", lhei, 
            "; lmat=\n")
        print(lmat)
    }
    dev.hold()
    on.exit(dev.flush())
    op <- par(no.readonly = TRUE)
    on.exit(par(op), add = TRUE)
    layout(lmat, widths = lwid, heights = lhei, respect = TRUE)
    if (!missing(RowSideColors)) {
        par(mar = c(margins[1L], 0, 0, 0.5))
        image(rbind(if (revC) 
            nr:1L
        else 1L:nr), col = RowSideColors[rowInd], axes = FALSE)
    }
    if (!missing(ColSideColors)) {
        par(mar = c(0.5, 0, 0, margins[2L]))
        image(cbind(1L:nc), col = ColSideColors[colInd], axes = FALSE)
    }
    par(mar = c(margins[1L], 0, 0, margins[2L]))
    if (!symm || scale != "none") 
        x <- t(x)
    if (revC) {
        iy <- nr:1
        if (doRdend) 
            ddr <- rev(ddr)
        x <- x[, iy]
        # switch rows
        my_org <- my_org[iy,]
    }
    else iy <- 1L:nr
    image(1L:nc, 1L:nr, x, xlim = 0.5 + c(0, nc), ylim = 0.5 + 
        c(0, nr), axes = FALSE, xlab = "", ylab = "", ...)
    #changed 1 to 3
    axis(3, 1L:nc, labels = labCol, las = 2, line = -0.5, tick = 0, 
        cex.axis = cexCol)
    if (!is.null(xlab)) 
        mtext(xlab, side = 3, line = margins[1L] )
    axis(2, iy, labels = labRow, las = 2, line = -0.5, tick = 0, 
        cex.axis = cexRow)
    
    # printing of values
    for (xcord in 1:ncol(my_org))
        for (ycord in 1:nrow(my_org))
            text(xcord, ycord, my_org[ycord,xcord])
            
    # try printing box around heatmap
    box(lty="solid", col="black")
    
    # try printing grid around values
    grid(nrow(my_grid), ncol(my_grid))
    
    #changed 4 to 2
    if (!is.null(ylab)) 
        mtext(ylab, side = 2, line = margins[2L] )
    if (!missing(add.expr)) 
        eval.parent(substitute(add.expr))
    par(mar = c(margins[1L], 0, 0, 0))
    if (doRdend) 
        plot(ddr, horiz = TRUE, axes = FALSE, yaxs = "i", leaflab = "none")
    else frame()
    par(mar = c(0, 0, if (!is.null(main)) 1 else 0, margins[2L]))
    if (doCdend) 
        plot(ddc, axes = FALSE, xaxs = "i", leaflab = "none")
    else if (!is.null(main)) 
        frame()
    if (!is.null(main)) {
        par(xpd = NA)
        title(main, cex.main = 1.5 * op[["cex.main"]])
    }
    invisible(list(rowInd = rowInd, colInd = colInd, Rowv = if (keep.dendro && 
        doRdend) ddr, Colv = if (keep.dendro && doCdend) ddc))
}


```

